---
title: "Problem Set 2"
author: "Andie Creel"
date: "2023-02-06"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vroom)
library(dplyr)
```
For both analyses this week, you will be using data from Mian and Sufi’s 2014 Econometrica article, What Explains the 2007-2009 Drop in Employment?. The analyses will not match the exact numbers in the paper, but the full replication set is available if you are interested in exploring it.

## Question One: Standard Errors
For this problem, use the dataset networth_delta_elas.csv, where county_fips is the county FIPS code, statename is the state FIPS code, elasticity is the Saiz elasticity measure, total is the number of households in each county, and netwp_h is the change in net worth within a county from 2006 to 2009.

### Write a function to esitmate the linear regression of networth change against a constant and the Saiz elasticiy. Report the coefficient on the elasticity.

Let $netwp$ be $Y_i$ and $elasticity$ be $X_i$ where $i$ indexes the county.

$$ Y_i = \beta_0 + \beta_1 X_i $$

When finding $\hat \beta$ by minimizing the squared error, 
$$\hat \beta= (X'X)^{-1} X'Y,$$
Note that $X$ is a matrix and the first column is equal to 1 for the constant.

```{r, message=FALSE}
# --------------------------------------------------------------
# read in data 
# --------------------------------------------------------------
# myData <- vroom('https://raw.githubusercontent.com/paulgp/applied-methods-phd/main/homework/Homework3/networth_delta_elas.csv')
# vroom_write(myData, file = "networth_delta_elas.csv")

myData <- vroom('networth_delta_elas.csv')

# --------------------------------------------------------------
# Function to estimate coefficients of OLS regressions
# --------------------------------------------------------------
myOLS <- function(Y, X){
  beta <- solve(t(X)%*%X)%*%t(X)%*%Y
  
  return(beta)
}

# --------------------------------------------------------------
# Estimating effect of elasticity on network change
# --------------------------------------------------------------
# getting matrices from data 
myX <- cbind(1, myData$elasticity) #constant and elasticity
myY <- myData$netwp_h

myBeta <-myOLS(Y = myY, X = myX)


# Checking
#lm(data = myData, formula = netwp_h ~ elasticity)
```

I estimate $\beta_0$ is `r myBeta[1]` and $\beta_1$ is `r myBeta[2]`.

### Next, estimate the homoskedastic SE, heteroskedasticity-robust SE, HC2, and HC3 standard errors for the elasticity estimate.

I use the code found on page 80 and 101 of *Econometrics* by Bruce Hansen (2014). 

```{r, message=FALSE}
# --------------------------------------------------------------
# Set UP
# --------------------------------------------------------------
x <- myX
y <- myY
e <- y - x%*%myBeta
#round(mean(e), 5)

n <- length(y)
k <- ncol(x)
a <- n/(n-k)
sig2 <- as.numeric((t(e) %*% e)/(n-k))
xx <- solve(t(x)%*%x) #X'X^-1

#Leverage 
leverage <- rowSums(x*(x%*%solve(t(x)%*%x)))

# --------------------------------------------------------------
# homoskedastic SE
# --------------------------------------------------------------
v0 <- xx*sig2
s0 <- sqrt(diag(v0)) # Homoskedastic formula

# check
# summary(lm(data = myData, formula = netwp_h ~ elasticity))

# --------------------------------------------------------------
# heteroskedasticity-robust SE
# --------------------------------------------------------------
u1 <- x*(e%*%matrix(1,1,k))
v1 <- n/(n-k)*xx %*% (t(u1)%*%u1) %*% xx
s1 <- sqrt(diag(v1)) # Heteroskedastic-robust (White formula)

# --------------------------------------------------------------
# HC2
# --------------------------------------------------------------
u2 <- x*((e/sqrt(1-leverage))%*%matrix(1,1,k))
v2 <- xx %*% (t(u2)%*%u2) %*% xx
s2 <- sqrt(diag(v2)) # HC2 (Horn-Horn-Duncan formula)

# --------------------------------------------------------------
# HC3
# --------------------------------------------------------------
u3 <- x*((e/(1-leverage))%*%matrix(1,1,k))
v3 <- xx %*% (t(u3)%*%u3) %*% xx
s3 <- sqrt(diag(v3)) # HC3 (Andrews formula)

```


Homoskedastic SE: $SE_0$ =  `r s0[1]`, $SE_1$ =  `r s0[2]`

Heteroskedasticity-Robust SE (aka Edgar Huber White formula): $SE_0$ =  `r s1[1]`, $SE_1$ =  `r s1[2]`

HC2 (aka Horn-Horn-Duncan formula) SE: $SE_0$ =  `r s2[1]`, $SE_1$ =  `r s2[2]`

HC3 (aka Andrews formula): $SE_0$ =  `r s3[1]`, $SE_1$ =  `r s3[2]`

### Now, we will estimate the three standard errors from Abadie et al. (2020) [see section 4]. I will walk you through the estimation.

$$V^{causal} = n^{-1} \Gamma^{-1}(\rho \Delta^{cond} + (1-\rho) \Delta^{ehw}) \Gamma^{-1}$$
$$V^{causal, sample} = \Gamma^{-1} \Delta^{cond}  \Gamma^{-1}$$
$$V^{descr} = n^{-1} (1 - \rho) \Gamma^{-1} \Delta^{ehw} \Gamma^{-1}$$
$$V^{ehw} = n^{-1} \Gamma^{-1} \Delta^{ehw} \Gamma^{-1}$$

$X$ is elasticity. $Z$ is our constant. $Y$ is the outcome of network change. 

```{r, message= FALSE}
# --------------------------------------------------------------
# Estimate e_hat as standard residual from Y ~ X + Z 
# --------------------------------------------------------------
n<- length(myData$elasticity)
myZ <- rep(1, n)
myX <- myData$elasticity 
myY <- myData$netwp_h

myBeta <- myOLS(myY, cbind(myZ,myX))

myStand_Resid <- myY - cbind(myZ,myX)%*%myBeta

# --------------------------------------------------------------
# Estimate the short regression of X~Z to calculate gamma_hat,
# which is the projeciton of X on Z. This is the mean when Z is a constant
# --------------------------------------------------------------
myGamma_i <- myOLS(myX, myZ)
myGamma_i <- as.numeric(myGamma_i)

# Check 
# mean(myX)

# --------------------------------------------------------------
# Estimate Gamma_hat
# --------------------------------------------------------------
Gamma_hat <- 1/n * sum((myX - myGamma_i*myZ)^2)

# --------------------------------------------------------------
# Estimate Delta_ehw 
# --------------------------------------------------------------
Delta_ehw <- 1/n * sum((myX - myGamma_i*myZ)*myStand_Resid^2*(myX - myGamma_i*myZ))

# --------------------------------------------------------------
# Now estimate V_EHW
# --------------------------------------------------------------
V_EHW <- (1/n)*Gamma_hat*Delta_ehw*Gamma_hat #ATTN: The check didn't match as close as I thought it would...
SE_EHW <- sqrt(V_EHW)

# --------------------------------------------------------------
# Estimate rho and V_descr
# --------------------------------------------------------------
myRho <- n/3006

V_descr <- (1-myRho)*V_EHW
SE_descr <- sqrt(V_descr)

# --------------------------------------------------------------
# Estimate G_hat
# --------------------------------------------------------------



```

* $\hat \epsilon_i$ is a vector
* I estimate $\hat \gamma$ = `r myGamma_i`
* I estimate $\hat \Gamma$ = `r Gamma_hat`
* I estimate $\hat \Delta^{ehw}$ = `r Delta_ehw`
* I estimate $\hat V^{ehw}$ = `r V_EHW`
* I estimate the standard error (descr) is `r SE_descr`. The EHW standard error was `r SE_EHW`, which is slightly larger. 
* 




# Citations
Hansen, Bruce. Econometrics. Princeton: Princeton University Press, 2014.
