---
title: "Problem Set Five"
author: "Andie Creel"
date: "2023-04-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)


library(vroom)
library(stargazer)
library(ivreg)
# library(ivpack)
library(dplyr)
```

# Question One (Comlier Share)

**Notes:**

Covariates in Angrist and Evans: age of mom, age of mom at first birth, race, ethnicity. 

My covariates: race and ethnicity

Samples: all women, married women, husbands of married women

## a) OLS regression 

$$y_i = \alpha +\beta x_i + \epsilon_i$$
where $y_i$ is some variable for labor supply and $x_i$ is $more_kids$. 



```{r, results='asis'}

#first subsample is only women with more than 2 kids 
myData_og <- vroom("ang_ev_1980.csv")

one <- lm(data = myData_og, formula = mom_worked ~ morekids +  blackm + hispm + whitem)
two <- lm(data = myData_og, formula = mom_weeks_worked ~ morekids +  blackm + hispm + whitem)

stargazer(one, two, type = "html", title = "OLS regression: effect of more than 2 kids")

```


## b) IV regression: instrument = same sex

```{r, results='asis'}
one.iv <- ivreg(data = myData_og, formula = mom_worked ~ morekids +  blackm + hispm + whitem | samesex + blackm + hispm + whitem)
two.iv <- ivreg(data = myData_og, formula = mom_weeks_worked ~ morekids +  blackm + hispm + whitem| samesex + blackm + hispm + whitem)

stargazer(one.iv, two.iv, type = "html", title = "IV regression: effect of more than 2 kids, instrumented with samesex")

```

This doesn't follow the paper's results very closely. 

#c) IV regression: instrument = twins of the same sex

```{r, results='asis'}
one.iv.2 <- ivreg(data = myData_og, formula = mom_worked ~ morekids +  blackm + hispm + whitem | twins_2 + blackm + hispm + whitem)
two.iv.2 <- ivreg(data = myData_og, formula = mom_weeks_worked ~ morekids +  blackm + hispm + whitem| twins_2 + blackm + hispm + whitem)

stargazer(one.iv.2, two.iv.2, type = "html", title = "OLS regression: effect of more than 2 kids")

```
This looks right. 


## d) Complier share for 2 instruments 

Complier share: 
$$E(D_i| Z_i = 1) - E(D_i| Z_i = 0)$$

**Same Sex:**

```{r}
mean(myData_og$morekids[myData_og$samesex == TRUE]) - mean(myData_og$morekids[myData_og$samesex == FALSE])
```


**Twins:**
```{r}
mean(myData_og$morekids[myData_og$twins_2 == TRUE]) - mean(myData_og$morekids[myData_og$twins_2 == FALSE])
```

Twins a stronger instrument. 

## e) Complier Share for different subsets

**Education greater than high school (moreths)**
```{r}
mean(myData_og$morekids[myData_og$twins_2 == TRUE & myData_og$moreths == TRUE]) - mean(myData_og$morekids[myData_og$twins_2 == FALSE & myData_og$moreths == TRUE])
```

**Black Mom**
```{r}
mean(myData_og$morekids[myData_og$twins_2 == TRUE & myData_og$blackm == TRUE]) - mean(myData_og$morekids[myData_og$twins_2 == FALSE & myData_og$blackm == TRUE])
```


**Hispanic Mom**
```{r}
mean(myData_og$morekids[myData_og$twins_2 == TRUE & myData_og$hispm == TRUE]) - mean(myData_og$morekids[myData_og$twins_2 == FALSE & myData_og$hispm == TRUE])
```

**White Mom**
```{r}
mean(myData_og$morekids[myData_og$twins_2 == TRUE & myData_og$whitem == TRUE]) - mean(myData_og$morekids[myData_og$twins_2 == FALSE & myData_og$whitem == TRUE])
```


## f) 
```{r}
# anderson.rubin.ci(one.iv, conflevel = .95)
```



# Question Two (Kitagawa Test)
I have only done this for twins as the IV. 


## a) 

```{r}
# P (treated by instrument)
df_P <- myData_og %>%
  filter(samesex == TRUE)

n1 <- nrow(df_P) %>% as.numeric()


sum_1_1 <- df_P %>%
  filter(mom_worked == T & morekids == T) %>%
  nrow()

sum_1_0 <- df_P %>%
  filter(mom_worked == T & morekids == F) %>%
  nrow()

sum_0_1 <- df_P %>%
  filter(mom_worked == F & morekids == T) %>%
  nrow()

sum_0_0 <- df_P %>%
  filter(mom_worked == F & morekids == F) %>%
  nrow()

P_y1_d1 <- n1^{-1} * sum_1_1
P_y1_d0 <- n1^{-1} * sum_1_0
P_y0_d1 <- n1^{-1} * sum_0_1
P_y0_d0 <- n1^{-1} * sum_0_0

#Reporting 
P_y1_d1 
P_y1_d0 
P_y0_d1 
P_y0_d0 


# Q (not treated by instrument)
df_Q <- myData_og %>%
  filter(samesex == F)

n0 <- nrow(df_Q) %>% as.numeric()
Q_y1_d1 <- n1^{-1} * sum(df_Q$mom_worked == TRUE & df_Q$morekids == TRUE)
Q_y1_d0 <- n1^{-1} * sum(df_Q$mom_worked == TRUE & df_Q$morekids == FALSE)
Q_y0_d1 <- n1^{-1} * sum(df_Q$mom_worked == FALSE & df_Q$morekids == TRUE)
Q_y0_d0 <- n1^{-1} * sum(df_Q$mom_worked == FALSE & df_Q$morekids == FALSE)

#Reporting 
Q_y1_d1 
Q_y1_d0 
Q_y0_d1 
Q_y0_d0 




```
Did this two different 

## b) 


```{r}
n <- n0 + n1
E <-  0.1
lambda <- n1/n0
sigma_y0_d0 <- ((1 - lambda)*P_y0_d0*(1 - P_y0_d0) + lambda*Q_y0_d0*(1 - Q_y0_d0))^.5
sigma_y1_d0 <- ((1 - lambda)*P_y1_d0*(1 - P_y1_d0) + lambda*Q_y1_d0*(1 - Q_y1_d0))^.5
sigma_y0_d1 <- ((1 - lambda)*P_y0_d1*(1 - P_y0_d1) + lambda*Q_y0_d1*(1 - Q_y0_d1))^.5
sigma_y1_d1 <- ((1 - lambda)*P_y1_d1*(1 - P_y1_d1) + lambda*Q_y1_d1*(1 - Q_y1_d1))^.5
sigma_y0_d0 
sigma_y1_d0 
sigma_y0_d1 
sigma_y1_d1 


T_stat <- (n0*n1/n)^.5 *max( max((Q_y0_d1 - P_y0_d1)/sigma_y0_d1, (Q_y1_d1 - P_y1_d1)/ sigma_y1_d1), 
                             max((Q_y0_d0 - P_y0_d0)/sigma_y0_d0, (Q_y1_d0 - P_y1_d0)/ sigma_y1_d0))

T_stat

```


## c) subsample 

```{r}
myData <- myData_og %>%
  mutate(H= lambda*((1 - mom_worked)*(1 - morekids)*P_y0_d0 +(mom_worked)*(1 - morekids)*P_y1_d0 +(1 - mom_worked)*(morekids)*P_y0_d1 +(mom_worked)*(morekids)*P_y1_d1) +
           (1 - lambda)*((1 - mom_worked)*(1 - morekids)*Q_y0_d0 +(mom_worked)*(1 - morekids)*Q_y1_d0 +(1 - mom_worked)*(morekids)*Q_y0_d1 +(mom_worked)*(morekids)*Q_y1_d1))

new_df <- myData[sample(1:nrow(myData), prob = myData$H, replace = T),]

```


```{r}
# P (treated by instrument)
df_P <- new_df %>%
  filter(samesex == TRUE)

n1 <- nrow(df_P) %>% as.numeric()


sum_1_1 <- df_P %>%
  filter(mom_worked == T & morekids == T) %>%
  nrow()

sum_1_0 <- df_P %>%
  filter(mom_worked == T & morekids == F) %>%
  nrow()

sum_0_1 <- df_P %>%
  filter(mom_worked == F & morekids == T) %>%
  nrow()

sum_0_0 <- df_P %>%
  filter(mom_worked == F & morekids == F) %>%
  nrow()

P_y1_d1 <- n1^{-1} * sum_1_1
P_y1_d0 <- n1^{-1} * sum_1_0
P_y0_d1 <- n1^{-1} * sum_0_1
P_y0_d0 <- n1^{-1} * sum_0_0
             


# Q (not treated by instrument)
df_Q <- new_df %>%
  filter(samesex == F)

n0 <- nrow(df_Q) %>% as.numeric()
Q_y1_d1 <- n1^{-1} * sum(df_Q$mom_worked == TRUE & df_Q$morekids == TRUE)
Q_y1_d0 <- n1^{-1} * sum(df_Q$mom_worked == TRUE & df_Q$morekids == FALSE)
Q_y0_d1 <- n1^{-1} * sum(df_Q$mom_worked == FALSE & df_Q$morekids == TRUE)
Q_y0_d0 <- n1^{-1} * sum(df_Q$mom_worked == FALSE & df_Q$morekids == FALSE)

```


```{r}
n <- n0 + n1
E <-  0.1
lambda <- n1/n0
sigma_y0_d0 <- ((1 - lambda)*P_y0_d0*(1 - P_y0_d0) + lambda*Q_y0_d0*(1 - Q_y0_d0))^.5
sigma_y1_d0 <- ((1 - lambda)*P_y1_d0*(1 - P_y1_d0) + lambda*Q_y1_d0*(1 - Q_y1_d0))^.5
sigma_y0_d1 <- ((1 - lambda)*P_y0_d1*(1 - P_y0_d1) + lambda*Q_y0_d1*(1 - Q_y0_d1))^.5
sigma_y1_d1 <- ((1 - lambda)*P_y1_d1*(1 - P_y1_d1) + lambda*Q_y1_d1*(1 - Q_y1_d1))^.5



T_stat <- (n0*n1/n)^.5 *max( max((Q_y0_d1 - P_y0_d1)/sigma_y0_d1, (Q_y1_d1 - P_y1_d1)/ sigma_y1_d1), 
                             max((Q_y0_d0 - P_y0_d0)/sigma_y0_d0, (Q_y1_d0 - P_y1_d0)/ sigma_y1_d0))

T_stat

```


## d) 
didn't complete 