---
title: "Problem Set 1"
subtitle: "Applied Methods"
author: "Andie Creel"
date: "2023-01-25"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vroom)
library(dplyr)

```

\newcommand{\indep}{\perp \!\!\! \perp}

## Question One: Randomization
This analysis will use the Dehijia and Wahba sample from the Lalonde dataset of the NSW experiment. The dataset is lalonde_nsw.csv. The outcome variable is re78 (real earnings in 1978). The treatment indicator is treat. The remaining variables are potential covariates. Assume for the purposes of this problem set that treat is completely randomly assigned.

### a. Calculate the average treatment effect of the policy $E(\tau_i)$ using a simple difference in means.

Under random assignment, we know that $\hat{\tau} = \sum_i \frac{Y_{i,1}}{D_i} - \sum_i \frac{Y_{i,0}}{1-D_i}$.

```{r, message=FALSE}
# --------------------------------------------------------------
# read in data 
# --------------------------------------------------------------
# myData <- vroom('https://raw.githubusercontent.com/paulgp/applied-methods-phd/main/homework/Homework1/lalonde_nsw.csv')
# vroom_write(myData, file = "lalonde_nsw.csv")

myData <- vroom('lalonde_nsw.csv')

# --------------------------------------------------------------
# calculating tau
# --------------------------------------------------------------
#calculating number of treated and control
D1 <- sum(myData$treat)
D0 <- sum(1-myData$treat)

# sum of outcome for treated 
Y1_df <- myData %>%
  filter(treat == 1) %>%
  select(re78)

Y1 <- sum(Y1_df$re78) ; rm(Y1_df)

# sum of outcome for control 
Y0_df <- myData %>%
  filter(treat == 0) %>%
  select(re78)

Y0 <- sum(Y0_df$re78) ; rm(Y0_df)

#calculating tau hat
t_hat <- Y1/D1 - Y0/D0

```

I find that $\hat{\tau}$ is `r round(t_hat, 2)`.

### b. Calculate the average treatment effect on the treated of the policy $E(\tau_i | treat = 1)$. How does it compare to part a?

Because treatment is randomly assigned (and we assume there are no violators to their treatment status), $E[Y_i(0) | D_i = 0] = E[Y_i(0) | D_i = 1]$. Therefore the average treatment on the treated with be the same as the average treatment effect and the answer does not differ from part a. 

### c. Test the null of $\tau_i = 0$ for all i using a randomization test. *N.B.* Hold fixed the number of treated and control (e.g. assume the treatment count would be held fixed) and permute the labels randomly 1000 times – you do not need to fully do every permutation (there would be too many). Report the quantile that your estimate from the previous question falls.

```{r}
# ----------------------------------------------------------------------------
# functions
# ----------------------------------------------------------------------------

#calculate the average treatement effect for a sample
calc_ATE <- function(df){

  #calculating number of treated and control
  D1 <- sum(df$treat_new)
  D0 <- sum(1-df$treat_new)
  
  # sum of outcome for treated 
  Y1_df <- df %>%
    filter(treat_new == 1) %>%
    select(re78)
  
  Y1 <- sum(Y1_df$re78) ; rm(Y1_df)
  
  # sum of outcome for control 
  Y0_df <- df %>%
    filter(treat_new == 0) %>%
    select(re78)
  
  Y0 <- sum(Y0_df$re78) ; rm(Y0_df)
  
  #calculating tau hat
  t_hat <- Y1/D1 - Y0/D0
  return(t_hat)
}

#get a sample ATE
get_sample_ate <- function(){
  # step one: randomize assignment of treatment and control
  mySample <- myData %>%
    mutate(treat_new = sample(treat, replace = FALSE))
 
  #step two: calculate t_hat 
  calc_ATE(mySample)
}


# ----------------------------------------------------------------------------
# Create a distribution and find p value
# ----------------------------------------------------------------------------

#create a distribution
myDist <- replicate(1000, get_sample_ate())

#indicate if sample's tau is bigger than estimated tau
myDist_df <- data.frame(sample_tau = myDist) %>%
  mutate(perc = if_else(sample_tau>t_hat, 1,0))

#calc p value
pval <- sum(myDist_df$perc)/1000

```
I find that the p-value is `r pval`. 


### d. Run a regression using robust standard errors (you may use canned software) of the outcome on the treatment dummy, and compare the p-values from this test to the previous answer.

```{r}
myReg <- lm(data = myData, formula = re78 ~ treat)
summary(myReg)

```
I find a p-value that is less than .01, as does the canned regression. We can also see that the average treatment effects are equivalent. 


## Question Two: Propensity Scores
This analysis will use the dataset from Problem 1 as well as the PSID dataset from Dehijia and Wahba, lalonde-psid.csv. These datasets have identical variables. The new dataset is a sample of observations from the Panel Survey of Income Dynamics that can be used as alternative control observations. Importantly, these observations were not in the initial randomization.

### a. Using age, education, hispanic, black, married, nodegree, RE74 and RE75, construct a propensity score using the treated group in lalonde-nsw.csv and the control sample of lalonde-psid.csv. Use a logit regression model to do so (you may use a canned routine to run the regression). Report the average p-score for the treated and control samples, and plot the propensity score densities for the treatment and control groups.

Step one: create new data set with treated from nsw and control for psid 

Step two: calculate propensity scores using logit 

$\pi(\mathbf{X_i}) = Pr(D_i = 1 | \mathbf{X_i})$

Step three: Subset by treated and control, calculate average propensity score for two groups. If there was random assignment, you would expect these to be the same (?)


```{r}
# mySample <- vroom("https://raw.githubusercontent.com/paulgp/applied-methods-phd/main/homework/Homework1/lalonde_psid.csv")
# vroom_write(mySample, "lalonde_psid.csv")

# ----------------------------------------------------------------------------
# read in data, make new data set
# ----------------------------------------------------------------------------
myData <- vroom("lalonde_nsw.csv") %>%
  filter(treat == 1)


mySample <- vroom("lalonde_psid.csv") %>%
  filter(treat == 0)

myData_2 <- bind_rows(myData, mySample) ; rm(myData, mySample)

myData_2 <- myData_2 %>% 
  select(treat, age, education, hispanic, black, married, nodegree, re74, re75)
  

# ----------------------------------------------------------------------------
# calculate pscores 
# ----------------------------------------------------------------------------

#logit
myLogit <- glm(treat ~ age + education + hispanic + black +married + nodegree + re74 + re75, data = myData_2, family = "binomial")

# predict treatment given model (pscore)
myData_2$p_score <- predict(myLogit, newdata = myData_2)

# ----------------------------------------------------------------------------
# avg pscore for treated and control 
# ----------------------------------------------------------------------------

# treated
treat_df <- myData_2 %>%
  filter(treat == 1)

p_score_treat <- mean(treat_df$p_score) ; rm(treat_df)

# control 
control_df <- myData_2 %>%
  filter(treat == 0)

p_score_control<- mean(control_df$p_score) ; rm(control_df)



```

Something is wrong, I am gretting a pscore of 
	
	
	


Assumptions: 
If we assume strong ignorability, $Y_i(0), Y_i(1) \indep D_i | \mathbf{X_i}$, then we know $(Y_i(0), Y_i(1) \indep D_i | \pi(\mathbf{X_i})$























```{r, eval=FALSE, include=FALSE}
# ----------------------------------------------------------------------------
# scratch 
# ----------------------------------------------------------------------------



get_sample_ate()

# try to do this with apply instead 

myDist <- data.frame(seq(1:1000))

sapply(myDist, get_sample_ate)

#initialization 
myDistribution <- data.frame(t_hat = seq(1:1000)) %>%
  mutate(t_hat = NA)

for (i in 1:1000) {
  
  # step one: randomize assignment of treatment and control
  mySample <- myData %>%
    mutate(treat_new = sample(treat, replace = FALSE))
 
   #step two: calculate t_hat and store
  myDistribution$t_hat[i] <- getATE(mySample)
  
}


```

