---
title: "Problem Set 1"
subtitle: "Applied Methods"
author: "Andie Creel"
date: "2023-01-25"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vroom)
library(dplyr)
# library(foreach)
# library(doParallel)
```

## Question One
Randomization. This analysis will use the Dehijia and Wahba sample from the Lalonde dataset of the NSW experiment. The dataset is lalonde_nsw.csv. The outcome variable is re78 (real earnings in 1978). The treatment indicator is treat. The remaining variables are potential covariates. Assume for the purposes of this problem set that treat is completely randomly assigned.

### a. Calculate the average treatment effect of the policy $E(\tau_i)$ using a simple difference in means.

Under random assignment, we know that $\hat{\tau} = \sum_i \frac{Y_{i,1}}{D_i} - \sum_i \frac{Y_{i,0}}{1-D_i}$.

```{r, message=FALSE}
# --------------------------------------------------------------
# read in data 
# --------------------------------------------------------------
# myData <- vroom('https://raw.githubusercontent.com/paulgp/applied-methods-phd/main/homework/Homework1/lalonde_nsw.csv')
# vroom_write(myData, file = "lalonde_nsw.csv")

myData <- vroom('lalonde_nsw.csv')

# --------------------------------------------------------------
# calculating tau
# --------------------------------------------------------------
#calculating number of treated and control
D1 <- sum(myData$treat)
D0 <- sum(1-myData$treat)

# sum of outcome for treated 
Y1_df <- myData %>%
  filter(treat == 1) %>%
  select(re78)

Y1 <- sum(Y1_df$re78) ; rm(Y1_df)

# sum of outcome for control 
Y0_df <- myData %>%
  filter(treat == 0) %>%
  select(re78)

Y0 <- sum(Y0_df$re78) ; rm(Y0_df)

#calculating tau hat
t_hat <- Y1/D1 - Y0/D0

```

I find that $\hat{\tau}$ is `r round(t_hat, 2)`.

### b. Calculate the average treatment effect on the treated of the policy $E(\tau_i | treat = 1)$. How does it compare to part a?

Because treatment is randomly assigned (and we assume there are no violators to their treatment status), $E[Y_i(0) | D_i = 0] = E[Y_i(0) | D_i = 1]$. Therefore the average treatment on the treated with be the same as the average treatment effect and the answer does not differ from part a. 

### c. Test the null of $\tau_i = 0$ for all i using a randomization test. *N.B.* Hold fixed the number of treated and control (e.g. assume the treatment count would be held fixed) and permute the labels randomly 1000 times – you do not need to fully do every permutation (there would be too many). Report the quantile that your estimate from the previous question falls.

```{r}
# ----------------------------------------------------------------------------
# functions
# ----------------------------------------------------------------------------

#calculate the average treatement effect for a sample
calc_ATE <- function(df){

  #calculating number of treated and control
  D1 <- sum(df$treat_new)
  D0 <- sum(1-df$treat_new)
  
  # sum of outcome for treated 
  Y1_df <- df %>%
    filter(treat_new == 1) %>%
    select(re78)
  
  Y1 <- sum(Y1_df$re78) ; rm(Y1_df)
  
  # sum of outcome for control 
  Y0_df <- df %>%
    filter(treat_new == 0) %>%
    select(re78)
  
  Y0 <- sum(Y0_df$re78) ; rm(Y0_df)
  
  #calculating tau hat
  t_hat <- Y1/D1 - Y0/D0
  return(t_hat)
}

#get a sample ATE
get_sample_ate <- function(){
  # step one: randomize assignment of treatment and control
  mySample <- myData %>%
    mutate(treat_new = sample(treat, replace = FALSE))
 
  #step two: calculate t_hat 
  calc_ATE(mySample)
}


# ----------------------------------------------------------------------------
# Create a distribution and find p value
# ----------------------------------------------------------------------------

#create a distribution
myDist <- replicate(1000, get_sample_ate())

#indicate if sample's tau is bigger than estimated tau
myDist_df <- data.frame(sample_tau = myDist) %>%
  mutate(perc = if_else(sample_tau>t_hat, 1,0))

#calc p value
pval <- sum(myDist_df$perc)/1000

```
I find that the p-value is `r pval`. 


### d. Run a regression using robust standard errors (you may use canned software) of the outcome on the treatment dummy, and compare the p-values from this test to the previous answer.

```{r}
myReg <- lm(data = myData, formula = re78 ~ treat)
summary(myReg)

```
I fid a p-value that is less than .01, as does the canned regression. We can also see that the average treatment effects are equivalent. 

## Question Two 














```{r, eval=FALSE, include=FALSE}
# ----------------------------------------------------------------------------
# scratch 
# ----------------------------------------------------------------------------



get_sample_ate()

# try to do this with apply instead 

myDist <- data.frame(seq(1:1000))

sapply(myDist, get_sample_ate)

#initialization 
myDistribution <- data.frame(t_hat = seq(1:1000)) %>%
  mutate(t_hat = NA)

for (i in 1:1000) {
  
  # step one: randomize assignment of treatment and control
  mySample <- myData %>%
    mutate(treat_new = sample(treat, replace = FALSE))
 
   #step two: calculate t_hat and store
  myDistribution$t_hat[i] <- getATE(mySample)
  
}




```

