---
title: "Homework Four"
author: "Andie Creel"
date: "2023-03-29"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(fixest)
library(vroom)
library(dplyr)
library(stargazer)
library(ggplot2)
```

# Diff-in-diff

10 time periods. 1000 ids. treatment turns on in period 5 (post). standard treatment and control. Be specific about what SE are used. 

```{r}
myData <- vroom("https://raw.githubusercontent.com/paulgp/applied-methods-phd/main/homework/dind_data.csv")


myData <- myData %>%
  mutate(time_id = as.factor(time_id)) %>%
  mutate(ids = as.factor(ids)) %>%
  mutate(treated_group = treated_group*1) %>%
  mutate(post = post*1)

```


## a) Estimate three regressions. 

```{r}

reg_1 <- feols(data = myData, fml = y_instant ~ post*treated_group, vcov = "hetero")
reg_2 <- feols(data = myData, fml = y_instant ~treated_group + i(post, treated_group) | time_id, vcov = "hetero")
reg_3 <- feols(data = myData, fml = y_instant ~  i(post, treated_group) | ids + time_id, vcov = "hetero")

etable(reg_1, reg_2, reg_3)
```

First, I'm using heteroskedastically robust standard errors. The point estimates are the same across regressions. The standard error decrease across regressions because more controls are added by using additional fixed effects.  

Note that time period one is being omitted so that we have a reference level. 

```{r}
myPlot <- myData %>%
  group_by(time_id, treated_group) %>%
  mutate(y_inst = mean(y_instant)) %>%
  select(time_id, treated_group, y_inst) %>%
  distinct() %>%
  mutate(time_id = as.numeric(time_id)) %>%
  mutate(treated_group = as.factor(treated_group))
  
ggplot(myPlot, aes(x = time_id, y = y_inst, color = treated_group)) + 
  geom_point() +
  theme_bw()
```

There is a steady decrease in Y across treatment and control. However, in the time period of treatment the treatment group didn't decrease. The decreasing trend continues after treatment but the gap is larger. 


## b) 

```{r}
myData_4 <- myData %>%
  mutate(time_id = relevel(time_id, '4'))

reg_4 <- feols(data = myData, fml = y_instant ~  i(time_id, treated_group, "4") | ids + time_id, vcov = "hetero")


etable(reg_4)

```



```{r}
estDF<- as.data.frame(reg_4$coeftable) %>%
  mutate(time_id = c(1,2, 3, 5, 6, 7, 8, 9, 10)) %>%
  rename(se = `Std. Error`)

ggplot(estDF, aes(x = time_id, y = Estimate)) + 
  geom_point() +
  geom_errorbar(aes(ymin=Estimate-se, ymax=Estimate+se), width=.2,
                 position=position_dodge(.9)) +
  geom_vline(xintercept = 4) +
  theme_bw()


```
Point est in T = 6: `r estDF$Estimate[5]`
Stand. error in T = 6: `r estDF$se[5]`

If we do not omit a level (such as period 4) then we do not have a comparison level and interpreting our treatment effect becomes non-sensible.  

## c) 

```{r}
mean(estDF$Estimate)

```
The treatment effect in part A is larger than that in part B. Part A uses a different reference level. 

