---
title: "Homework Four"
author: "Andie Creel"
date: "2023-03-29"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(fixest)
library(vroom)
library(dplyr)
library(stargazer)
library(ggplot2)
library(stats)
```

# Diff-in-diff

10 time periods. 1000 ids. treatment turns on in period 5 (post). standard treatment and control. Be specific about what SE are used. 

```{r}
myData_og <- vroom("https://raw.githubusercontent.com/paulgp/applied-methods-phd/main/homework/dind_data.csv")


myData_og <- myData_og %>%
  mutate(time_id = as.factor(time_id)) %>%
  mutate(ids = as.factor(ids)) %>%
  mutate(treated_group = treated_group*1) %>%
  mutate(post = post*1)

```


## a) Estimate three regressions. 

```{r}
myData <- myData_og %>%
  mutate(time_id = relevel(time_id, ref = "1"))


reg_1 <- feols(y_instant ~ treated_group * post, data = myData, vcov = "hetero")
reg_2 <- feols(y_instant ~  treated_group + treated_group*post | time_id, data = myData, vcov = "hetero")
reg_3 <- feols(y_instant ~  treated_group*post | time_id + ids, data = myData, vcov = "hetero")

etable(reg_1, reg_2, reg_3)
```

First, I'm using heteroskedastically robust standard errors. The point estimates are the same across regressions. The standard error decrease across regressions because more controls are added by using additional fixed effects.  

Note that time period one is being omitted so that we have a reference level. 

```{r}
myPlot <- myData %>%
  group_by(time_id, treated_group) %>%
  mutate(y_inst = mean(y_instant)) %>%
  select(time_id, treated_group, y_inst) %>%
  distinct() %>%
  mutate(time_id = as.numeric(time_id)) %>%
  mutate(treated_group = as.factor(treated_group))
  
ggplot(myPlot, aes(x = time_id, y = y_inst, color = treated_group)) + 
  geom_point() +
  theme_bw() +
  ggtitle("Instant")
```

There is a steady decrease in Y across treatment and control. However, in the time period of treatment the treatment group didn't decrease. The decreasing trend continues after treatment but the gap is larger. 


## b) 

```{r}

# regression 
myData <- myData_og %>%
  mutate(time_id = relevel(time_id, ref = "4"))

reg_4 <- feols(y_instant ~  treated_group*time_id | time_id + ids, data = myData, vcov = "hetero")

etable(reg_4)

# plot 
estDF_4<- as.data.frame(reg_4$coeftable) %>%
  mutate(time_id = c(1,2, 3,  5, 6, 7, 8, 9, 10)) %>%
  rename(se = `Std. Error`)

ggplot(estDF_4, aes(x = time_id, y = Estimate)) + 
  geom_point() +
  geom_errorbar(aes(ymin=Estimate-se, ymax=Estimate+se), width=.2,
                 position=position_dodge(.9)) +
  geom_vline(xintercept = 4) +
  theme_bw() +
  ggtitle("Instant, ref is 4")


```
Point est in T = 6: `r estDF_4$Estimate[5]`
Stand. error in T = 6: `r estDF_4$se[5]`

If we do not omit a level (such as period 4) then we do not have a comparison level and interpreting our treatment effect becomes non-sensible.  

## c) 

$$E[y(1) - y(0)]$$

```{r}

y_0 <- estDF_4 %>%
  filter(time_id <= 4)

y_1 <- estDF_4 %>%
  filter(time_id > 4)


mean(y_1$Estimate) - mean(y_0$Estimate)

```
They're the same. 


## d) 

```{r}
# regression 
myData <- myData_og %>%
  mutate(time_id = relevel(time_id, ref = "3"))

reg_5 <- feols(y_instant ~  treated_group*time_id | time_id + ids, data = myData, vcov = "hetero")

etable(reg_5)

# plot 
estDF_5<- as.data.frame(reg_5$coeftable) %>%
  mutate(time_id = c(1,2, 4,  5, 6, 7, 8, 9, 10)) %>%
  rename(se = `Std. Error`)

ggplot(estDF_5, aes(x = time_id, y = Estimate)) + 
  geom_point() +
  geom_errorbar(aes(ymin=Estimate-se, ymax=Estimate+se), width=.2,
                 position=position_dodge(.9)) +
  geom_vline(xintercept = 3) +
  theme_bw() +
  ggtitle("Instant, ref is 3")


```


## e) 

```{r}

myData <- myData_og %>%
  mutate(time_id = relevel(time_id, ref = "1"))

reg_1.e<- feols(y_dynamic ~ treated_group * post, data = myData, vcov = "hetero")
reg_2.e<- feols(y_dynamic ~  treated_group + treated_group*post | time_id, data = myData, vcov = "hetero")
reg_3.e<- feols(y_dynamic ~  treated_group*post | time_id + ids, data = myData, vcov = "hetero")

etable(reg_1.e, reg_2.e, reg_3.e)
```
The treatment effect has increased.

```{r}
myPlot <- myData %>%
  group_by(time_id, treated_group) %>%
  mutate(y_dyn = mean(y_dynamic)) %>%
  select(time_id, treated_group, y_dyn) %>%
  distinct() %>%
  mutate(time_id = as.numeric(time_id)) %>%
  mutate(treated_group = as.factor(treated_group))
  
ggplot(myPlot, aes(x = time_id, y = y_dyn, color = treated_group)) + 
  geom_point() +
  theme_bw() +
  ggtitle("Dynamic ")
```
We see that the decreasing trend is now halted for 3 time periods for the treated group (6, 7, 8).

```{r}

# regression 
myData <- myData_og %>%
  mutate(time_id = relevel(time_id, ref = "4"))

reg_4.e <- feols(y_dynamic ~  treated_group*time_id | time_id + ids, data = myData, vcov = "hetero")

etable(reg_4.e)

# plot 
estDF_4.e<- as.data.frame(reg_4.e$coeftable) %>%
  mutate(time_id = c(1,2, 3, 5, 6, 7, 8, 9, 10)) %>%
  rename(se = `Std. Error`)

ggplot(estDF_4.e, aes(x = time_id, y = Estimate)) + 
  geom_point() +
  geom_errorbar(aes(ymin=Estimate-se, ymax=Estimate+se), width=.2,
                 position=position_dodge(.9)) +
  geom_vline(xintercept = 4) +
  theme_bw() +
  ggtitle("Dynamic, ref is 4 ")


```
Rather than there being one time period where the trend is forgone, there are multiple. Therefore the treatment effect for each time period increase from time period 5 through 8, and then the treatment affect stabilizes for the remaining time periods. 

## f) 

```{r}
myData <- myData_og %>%
  mutate(time_id = relevel(time_id, ref = "1"))

myPlot <- myData %>%
  group_by(time_id, treated_group) %>%
  mutate(y_dyn = mean(y_dynamic2)) %>%
  select(time_id, treated_group, y_dyn) %>%
  distinct() %>%
  mutate(time_id = as.numeric(time_id)) %>%
  mutate(treated_group = as.factor(treated_group))
  
ggplot(myPlot, aes(x = time_id, y = y_dyn, color = treated_group)) + 
  geom_point() +
  theme_bw() +
  ggtitle("Dynamic 2")
```
The pre-trend is not perfectly parallel. However, it this were empirical data I'd argue that it fits well enough to do a diff-in-diff. 
```{r}
# regression 
myData <- myData_og %>%
  mutate(time_id = relevel(time_id, ref = "4"))

reg_4.f <- feols(y_dynamic2 ~  treated_group*time_id | time_id + ids, data = myData, vcov = "hetero")

etable(reg_4.f)

# plot 
estDF_4.f<- as.data.frame(reg_4.f$coeftable) %>%
  mutate(time_id = c(1,2, 3,  5, 6, 7, 8, 9, 10)) %>%
  rename(se = `Std. Error`)

ggplot(estDF_4.f, aes(x = time_id, y = Estimate)) + 
  geom_point() +
  geom_errorbar(aes(ymin=Estimate-se, ymax=Estimate+se), width=.2,
                 position=position_dodge(.9)) +
  geom_vline(xintercept = 4) +
  theme_bw() +
  ggtitle("Dynamic 2, ref is 4")


```

```{r}
# regression 
myData <- myData_og %>%
  mutate(time_id = relevel(time_id, ref = "3"))

reg_5.f <- feols(y_dynamic2 ~  treated_group*time_id | time_id + ids, data = myData, vcov = "hetero")

etable(reg_5.f)

# plot 
estDF_5.f<- as.data.frame(reg_5.f$coeftable) %>%
  mutate(time_id = c(1,2, 4,  5, 6, 7, 8, 9, 10)) %>%
  rename(se = `Std. Error`)

ggplot(estDF_5.f, aes(x = time_id, y = Estimate)) + 
  geom_point() +
  geom_errorbar(aes(ymin=Estimate-se, ymax=Estimate+se), width=.2,
                 position=position_dodge(.9)) +
  geom_vline(xintercept = 3) +
  theme_bw() +
  ggtitle("Dynamic 2, ref is 3")


```

## g) 
In part A I already used heteroskedastically robust standard errors. I now repeat part A with robust SE that are clustered by id. 

```{r}
myData <- myData_og %>%
  mutate(time_id = relevel(time_id, ref = "1"))

reg_3.g.hom <- feols(y_instant ~  treated_group*post | time_id + ids, data = myData, vcov = "iid")
reg_3.g.rob <- feols(y_instant ~  treated_group*post | time_id + ids, data = myData, vcov = "hetero")
reg_3.g.clust <- feols(y_instant ~  treated_group*post | time_id + ids, data = myData, cluster = myData$ids)


etable(reg_3.g.hom, reg_3.g.rob, reg_3.g.clust)
```
The robust standard error is smaller than the clustered standard error, but larger than than the SE when assuming errors are IID. 

```{r}
# plotting all the graphs in a grid
par(mfrow = c(2, 5)) 

#Auto corrleation graphs
auto_corr <- myData_og %>%
  mutate(residuals = resid(reg_3.g.rob)) %>%
  group_by(time_id) %>%
  summarise(cor=list(acf(residuals, lag.max = 1000)))


```

There doesn't appear to be any auto correlation within groups to be concerned about. 

# Time Series 

```{r}
rm(list = ls())

# reading in data 
myData_og <- vroom("https://raw.githubusercontent.com/paulgp/applied-methods-phd/main/homework/Homework6/data/hrs_eventdata.csv")

# hhidpn: individual 
# wave: wave of survey / time index of study
# wave_hosp: time of event
# oop spend: out of pocket spending 

```


## a) 
wave_hosp corresponds with $E_i$ in our study. 

```{r}
# constructing E
myData <- myData_og %>%
  mutate(D = if_else(wave_hosp <= wave, 1, 0))

# share treated in each time period 
myData %>% 
  group_by(wave) %>%
  mutate(share_treated = sum(D)/length(D)) %>%
  select(share_treated) %>%
  distinct()


```

## b) 

```{r}
myReg_1 <- feols(oop_spend ~ D | wave + hhidpn, data = myData)

etable(myReg_1)

```


Interpretation of inference given panel setting: 

## c) 
```{r}
# Set level as last time period
myData <- myData_og %>%
  mutate(wave_hosp = as.factor(wave_hosp)) %>%
  mutate(wave_hosp = relevel(wave_hosp, ref = "11"))

# YOU ARE HERE 

```



